Relative Stable Maps in Genus One via Central Alignments

### 2019-01-31 GenusOneRelativeMaps.tex
Finished proof that M_v^tilde integral

\subsection{Splitting and gluing punctured logarithmic maps}\label{subsection punctured maps} We still start by considering the usual moduli space $\Mcal$ of log stable maps. In \cite{PuncturedMaps}, a recursive description of the logarithmic strata of $\Mcal$ is given, in terms of moduli spaces of punctured maps. As discussed above, the logarithmic strata $\Scal$ of $\Mcal$ are indexed by combinatorial types of tropical curves. Given such a combinatorial type $\Delta$, each vertex $v$ of the associated graph defines a moduli space of punctured stable maps:
\begin{equation*} \ol\Mcal_v = \Mpunct{g(v)}{\alpha(v)}{Z_v}{d(v)}. \end{equation*}
Here $g(v)$ and $d(v)$ are the genus and degree assigned to the vertex, while $Z_v$ is the logarithmic stratum of the target corresponding to the tropical stratum into which the vertex is mapped by the combinatorial type (in our setting, $Z_v=\PP^N$ or $H$). The length of the vector $\alpha(v)$ is given by the number of edges and legs adjacent to $v$, and the entries (which represent tangencies, and may be positive or negative) are given by the expansion factors along those edges and legs. If all of the tangencies are positive, then this is nothing but an ordinary moduli space of log stable maps.

The corresponding logarithmic stratum is then obtained as a follows. For each vertex $v$ and each adjacent edge $E \ni v$ we define $\widetilde\Mcal_v^E$ to be the stack $\ol\Mcal_v$ equipped with the log structure
\begin{equation*} p_{v,E}^\st \Mcal_{\Ccal_v^\circ} \end{equation*}
where $p_{v,E} \colon \ol\Mcal_v \to \Ccal_v^\circ$ is the marking section corresponding to $E$. This definition makes $p_{v,E}$ into a strict logarithmic map, and composing with the universal map produces a logarithmic evaluation morphism $\ev_{v,E} \colon \widetilde\Mcal_v^E \to Z_E$, where $Z_E \subseteq Z_v$ is the logarithmic stratum into which the marking corresponding to $E$ is mapped.

If $E_1,\ldots,E_k$ are the edges adjacent to $v$ then we define
\begin{equation*} \widetilde\Mcal_v = \widetilde\Mcal_v^{E_1} \times_{\ol\Mcal_v} \ldots \times_{\ol\Mcal_v} \widetilde\Mcal_v^{E_k} \end{equation*}
where the fibre product is taken in the fs log category. With this definition, all the evaluation maps $\ev_{v,E} \colon \widetilde{\Mcal}_v \to Z_E$ are naturally logarithmic morphisms. Then the corresponding logarithmic stratum is obtained by taking the following fibre product (again, in the fs log category):
\bcd
\Scal \ar[r] \ar[d] \ar[rd,phantom,"\square"] & \prod_v \widetilde\Mcal_v \ar[d] \\
\prod_{E} Z_E \ar[r] & \prod_{E} Z_E \times Z_E.
\ecd
In the above discussion, there were two steps at which we needed to take fibre products in the fs log category: first, when constructing each $\widetilde{\Mcal}_v$ and second, when construction $\Scal$. Typically the fs fibre product does not agree with the usual fibre product of stacks; rather, one must take the fibre product in the category of coherent log stacks (which does agree with the usual fibre product) and then integralise and saturate the result.

We will now argue that, in our situation, the fibre products of coherent log stacks are automatically integral. This means that the integralisation process is redundant, and so the fs fibre product is obtained from the ordinary fibre product by a process of saturation, which amounts to taking a finite cover of the underlying stack (whose degree we will compute).

\begin{lemma}

\end{lemma}

>>>>

\subsection{Splitting and gluing punctured logarithmic maps}\label{subsection punctured maps} We still start by considering the usual moduli space $\Mcal$ of log stable maps. In \cite{PuncturedMaps}, a recursive description of the logarithmic strata of $\Mcal$ is given, in terms of moduli spaces of punctured maps. As discussed above, the logarithmic strata $\Scal$ of $\Mcal$ are indexed by combinatorial types of tropical curves. Given such a combinatorial type $\Delta$, each vertex $v$ of the associated graph defines a moduli space of punctured stable maps:
\begin{equation*} \ol\Mcal_v = \Mpunct{g(v)}{\alpha(v)}{Z_v}{d(v)}. \end{equation*}
Here $g(v)$ and $d(v)$ are the genus and degree assigned to the vertex, while $Z_v$ is the logarithmic stratum of the target corresponding to the tropical stratum into which the vertex is mapped by the combinatorial type (in our setting, $Z_v=\PP^N$ or $H$). The length of the vector $\alpha(v)$ is given by the number of flags adjacent to $v$, and the entries (which represent tangencies, and may be positive or negative) are given by the expansion factors along those flags. If all of the tangencies are positive, then this is nothing but an ordinary moduli space of log stable maps.

The corresponding logarithmic stratum is then obtained as a follows. For each vertex $v$ and each adjacent flag $E \ni v$ we define $\widetilde\Mcal_v^E$ to be the stack $\ol\Mcal_v$ equipped with the log structure
\begin{equation*} p_{v,E}^\st \Mcal_{\Ccal_v^\circ} \end{equation*}
where $p_{v,E} \colon \ol\Mcal_v \to \Ccal_v^\circ$ is the marking section corresponding to $E$. This definition makes $p_{v,E}$ into a strict logarithmic map, and composing with the universal map produces a logarithmic evaluation morphism $\ev_{v,E} \colon \widetilde\Mcal_v^E \to Z_E$, where $Z_E \subseteq Z_v$ is the logarithmic stratum into which the marking corresponding to $E$ is mapped.

If $E_1,\ldots,E_k$ are the flags adjacent to $v$ then we define
\begin{equation*} \widetilde\Mcal_v = \widetilde\Mcal_v^{E_1} \times_{\ol\Mcal_v} \ldots \times_{\ol\Mcal_v} \widetilde\Mcal_v^{E_k} \end{equation*}
where the fibre product is taken in the fs log category. With this definition, all the evaluation maps $\ev_{v,E} \colon \widetilde{\Mcal}_v \to Z_E$ are naturally logarithmic morphisms. Then the corresponding logarithmic stratum is obtained by taking the following fibre product (again, in the fs log category):
\bcd
\Scal \ar[r] \ar[d] \ar[rd,phantom,"\square"] & \prod_v \widetilde\Mcal_v \ar[d] \\
\prod_{E} Z_E \ar[r] & \prod_{E} Z_E \times Z_E.
\ecd
In the above discussion, there were two steps at which we needed to take fs fibre products: first, when constructing each $\widetilde{\Mcal}_v$ and second, when construction $\Scal$. Typically the fs fibre product does not agree with the usual fibre product of stacks; rather, one must take the fibre product in the category of coherent log stacks (which \emph{does} agree with the usual fibre product) and then integralise and saturate the result.

We will now argue that in our situation, when we take fibre products of coherent log stacks, the result is automatically integral. This means that the integralisation process is redundant, and so the fs fibre product is obtained from the ordinary fibre product by a process of saturation. This amounts to taking a finite cover of the underlying stack (whose degree we will compute).

The following lemma holds for arbitrary moduli spaces of punctured maps.

\begin{lemma} Given a combinatorial type $\Delta$ indexing a stratum of $\Mcal$ and a vertex $v$ of the associated graph, consider the corresponding moduli space of punctured stable maps $\ol\Mcal_v$. As in the above discussion, define for each adjacent flag $E \ni v$ the log stack $\widetilde\Mcal_v^E$ by pulling back the log structure of the punctured curve along the marking section. Consider the fibre product \emph{in the category of coherent log stacks}:
\begin{equation*} \widetilde\Mcal^{\operatorname{not-fs}}_v = \widetilde{\Mcal}_v^{E_1} \times_{\ol\Mcal_v} \ldots \times_{\ol\Mcal_v} \widetilde{\Mcal}_v^{E_k}. \end{equation*}
Then $\widetilde\Mcal^{\operatorname{not-fs}}_v$ is integral (and consequently, the map $\widetilde\Mcal_v \to \widetilde\Mcal^{\operatorname{not-fs}}_v$ is finite).
\end{lemma}
\begin{proof} Locally we have a chart for $\ol\Mcal_v$ given by $Q$ (the minimal monoid as constructed, for instance, in \cite[Construction 1.16]{GrossSiebertLog}) and charts for $\widetilde\Mcal_v^{E_i}$ given by $Q_{E_i}$ where $Q_{E_i} = Q \oplus \N$ if $E_i$ corresponds to an ordinary marked point, and $Q_{E_i}$ is the submonoid of $Q \oplus \Z$ generated by $Q \oplus \N$ and $(\ol\varphi_{\eta_i}(1),u_{E_i})$ if $E_i$ corresponds to a puncturing (so that $u_{E_i} \in \Z_{<0}$). In either case, we have $Q_{E_i} \subseteq Q \oplus \Z \subseteq Q^{\operatorname{gp}} \oplus \Z$ (since $Q$ is integral) and so $Q_{E_i}$ is integral (recall that a monoid is integral if and only if it admits an injective homomorphism into a group). In this notation, $\widetilde\Mcal_v^{\operatorname{not-fs}}$ admits a chart given by the monoid:
\begin{equation*} P = Q_{E_1} \oplus_Q Q_{E_2} \oplus_Q \ldots \oplus_Q Q_{E_k}.\end{equation*}
If we consider the map $D \colon Q^{r-1} \to Q^r$ given by the $(r-1) \times r$ matrix
\begin{equation*} D= \begin{pmatrix} 1 & 1 & 0 & 0 & \ldots & 0 \\
0 & 1 & 1 & 0 & \ldots & 0 \\
\, & \, & \, &\vdots & \, & \, \\
0 & 0 & \ldots & 0 & 1 & 1 \end{pmatrix}\end{equation*}
then the monoid $P$ can be expressed as a cokernel
\begin{equation*} Q^{r-1} \to \oplus_{i=1}^r Q_{E_i} \to P \end{equation*}
where the first map is obtained by composing $D$ with the natural inclusion $Q^r \to \oplus_{i=1}^r Q_{E_i}$. We may now form a diagram
\bcd
Q^{r-1} \ar[d,"\Id"] \ar[r] & \oplus_{i=1}^r Q_{E_i} \ar[r] \ar[d,hook] & P \ar[d] \\
Q^{r-1} \ar[r] & Q^r \oplus \Z^r \ar[r] & S \oplus \Z^r
\ecd
where both rows are cokernel diagrams and the map $P \to S\oplus\Z^r$ is induced by the universal property. Here $S$ is the cokernel of $D$ and is given by:
\begin{equation*} S = \begin{cases} Q \qquad \text{if } r=1 \\ Q^{\gp} \qquad \text{if } r \geq 2 \end{cases}.\end{equation*}
In any case, $S$ is integral and the map $P \to S \oplus \Z^r$ is injective because the middle vertical maps in injective. Hence we conclude that $P$ is integral, as required. \end{proof}

The previous lemma holds for arbitrary moduli spaces of punctured maps. The next lemma, on the other hand, is special to our setting.

\begin{lemma} Consider the fibre product in the category of coherent log stacks:
\bcd
\Scal^{\operatorname{not-fs}} \ar[r] \ar[d] \ar[rd,phantom,"\square"] & \prod_v \widetilde\Mcal_v \ar[d] \\
\prod_{E} Z_E \ar[r] & \prod_{E} Z_E \times Z_E.
\ecd
Then $\Scal^{\operatorname{not-fs}}$ is integral.\end{lemma}
